---
title: "R-Assignment-BCB546-Spring2021-Topping"
author: "Nick Topping"
date: "3/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Assignment

To start, create a repository on GitHub.
My repository is named R-Assignment-BCB546-Spring2021-Topping.
To load the working repository into R Studio:

1. Clone the repository to your clipboard
2. Open R Studio and click File, New Project, Version Control, Git
3. Paste the repository URL, fill in the directory name and click Create Project

Next, open both data files, `fang_et_al_genotypes` and `snp_position.txt`.
Both of these files were pulled from my computer where I already had them saved
``` {r}
fang_et_al_genotypes <- read.table("https://raw.githubusercontent.com/EEOB-BioData/BCB546-Spring2021/master/assignments/UNIX_Assignment/fang_et_al_genotypes.txt", header = TRUE,)
snp_position <- read.table("https://raw.githubusercontent.com/EEOB-BioData/BCB546-Spring2021/master/assignments/UNIX_Assignment/snp_position.txt", header = TRUE, sep = "\t")
```
To quickly look at the data files, double click on them in the Environment. This will show the data in a .txt file that you can see.

Note: You can also use the function `view(fang_et_al_genotypes)` will also show the data, but it isn't working in my markdown file for some reason...

Below are a few codes you can use to inspect the data

`object.size()` shows the file size of the data frame

`nrow()` shows the number of rows

`ncol()` shows the number of columns

`dim()` shows the dimensions of the data frame by row and column

`head()` show first 6 rows

`tail()` shows last 6 rows

`str()` shows the structure of the data frame: the name, type and preview of data in each column

`names()` shows the names attributes for the data frame, which are also the column names

the function `colnames()` also will show the column names

`sapply(dataframe, class)` shows the class of each column of data

## fang_et_al_genotypes

``` {r}
object.size(fang_et_al_genotypes)
```

``` {r}
nrow(fang_et_al_genotypes)
```

``` {r}
ncol(fang_et_al_genotypes)
```

``` {r}
dim(fang_et_al_genotypes)
```

Note: The output for functions `head()` `tail()` `str()` `names()` and `sapply(dataframe, class)` are not included in this markdown file for this data table as these outputs are very long and clutter the console

## snp_position

``` {r}
nrow(snp_position)
```

``` {r}
ncol(snp_position)
```

``` {r}
dim(snp_position)
```

``` {r}
str(snp_position)
```

``` {r}
names(snp_position)
```

``` {r}
sapply(snp_position, class)
```

For this data table, only `head()` and `tails()` were left out of the snp_position analyses


